Understanding domestic abuse

### 2019-07-29 case_for_support_TM_hacked.tex
Anna cfs update

This research programme consists of four research topics aiming to deepen our understanding of domestic abuse victimisation, by exploiting the rich information contained in the CSEW, and combine it with insights from other data sources. First, we will explore the demographic and socio-economic characteristics of victims of domestic abuse, and the predictors of suffering serious harm. Second, we will examine the reasons behind not just under-reporting from victims, but also police mis-recording of domestic abuse cases, and victim's perception of the police after reporting an incident. Third, we will attempt to quantify the long-term effects of being a victim of domestic abuse, and witnessing domestic abuse at home as a child. Finally, we explore the association between exogenous, time-varying factors, such as changes in alcohol consumption and financial resilience, and the prevalence of domestic abuse.

@@

The research projects detailed below draw on various data sources. All of the projects rely on the CSEW, an annual, cross-sectional representative survey collecting information on victims of crime across England and Wales, including 30,000--40,000 households every year \cite{OfficeforNationalStatistics2019}. The survey has a self-completion module containing questions about the respondent's experiences of domestic abuse throughout their life (since the age of 16), and in the past 12 months in particular. Given the serious problem of underreporting, one of the unique characteristic of domestic abuse, the CSEW is the most reliable source of information on the prevalence of domestic abuse in England and Wales. We have access to this dataset via the UK Data Service.

@@

These two datasets have different merits. The CSEW provides us with rich demographic and socio-economic information on the respondent, and to investigate the casual effects of domestic abuse victimisation, we can construct a control group from those individuals who have not experienced domestic abuse. However, it does not contain any information on the perpetrators, and has limited longitudinal information. In contrast, the WMP dataset has information on both the victim and the offender, and records all reported incidents for the same victim-offender pair, allowing us to analyse temporal patterns, and identify high-risk times of the year. We will not link the individuals in these datasets at any time.

@@

This research programme addresses four overarching topics, exploring different aspects of domestic abuse. Drawing on the extensive, rich respondent information in the CSEW, and the unique, temporal nature of the WMP data, this research complements, extends and deepens our understanding of domestic abuse. In outlining the analysis plan for each of these research topics, we will specify a set of statistical analyses we plan to conduct. We will preregister all of our analysis plans and complement all analyses with robustness checks, and further exploratory regressions as the data allows (see, for an example of this approach in some of our \href{https://cog.warwick.ac.uk/~pssmar/Trendl_Stewart_Mullett.pdf}{prior work} on domestic abuse).

@@

Investigating the explanatory power of a broader set of victim characteristics within one (logistic) regression model will provide us with a deeper understanding of the relative importance of these factors in predicting domestic abuse victimisation, compared to a purely descriptive statistics approach. We are also interested in how the predictive power of victim characteristics vary with the type of abuse suffered (physical abuse, threats, sexual abuse).

@@

We will also use the CSEW to explore the police's response to reported cases of domestic abuse. Increasing victim's trust in the Criminal Justice System (CJS) through improving law enforcement response to reported incidents is key for improving willingness to report. Mis-recording of domestic abuse is a serious problem across police forces, for example, a recent inspection reported that WMP have failed to record 25\% of reported crimes that were domestic abuse-related \cite{HerMajestysInspectorateofConstabularyandFires&RescueServices2019}. In this research topic, we will explore whether the extent of crime mis-recording depends on victim characteristics. Using the CSEW, we will be able to tell whether the type of domestic abuse reported by the respondent is likely to amount to a crime according to the list of notifiable offences \cite{countingrules}. A multinomial regression approach will allow us to examine the extent to which police mis-reporting is different across broad crime types (sexual, physical violence, harassment and stalking), and depends upon the socio-economic characteristics of the victim and police forces (e.g., funding per police officer, deprivation). Complementing this analysis with a random forest approach will reveal the role of interactions between individual and police force characteristics. Lastly, these analyses can also be applied to examine the effect upon victim satisfaction with the police, and their perception of their personal security after reporting a crime.

@@

 Domestic abuse has long-lasting adverse effects on victims and children witnessing the abuse. In this research topic, our is aim to explore some of these consequences. The 2016 CSEW included a module asking respondents about whether they witnessed domestic abuse at home as a child. Descriptive statistics of this dataset revealed that those who reported to have witnessed abuse as a child were significantly more likely to report long-term health problems, and live in a single parent household as an adult \cite{ONSChildhood}. First, we will consolidate and extend these findings using propensity score matching, a statistical approach that allows for causal inferences to be drawn about the effect of witnessing abuse, whilst also controlling for a range of confounds, such as economic deprivation. 
 %Gaining a deeper understanding of the tangible, far-reaching consequences of domestic abuse will help to design survivor support programmes and quantify the societal harm caused by it.

@@

We will further complement this analysis using WMP data, by identifying whether younger offenders' home addresses can be linked to a past domestic abuse incident. It is not possible to predict whether witnessing domestic abuse increases the chance of becoming a young offender, but we can examine whether it predicts the type of crime they engage in. For example, whether, after controlling for the socio-economic characteristics of the neighbourhood (using the IMD), such individuals are more likely to commit similarly violent acts than other young offenders.

 These analyses will consist the first UK-based comprehensive quantitative exploration of the causal effect of domestic abuse and childhood abuse on socio-economic outcomes. The resulting paper will be targeted at The Lancet Public Health, and Social Science and Medicine.

@@

In addition, drawing on previous findings about the link between financial stress and domestic abuse, we will explore what effect the roll-out of the Universal Credit (UC) across the West Midlands had on the reported number of domestic abuse cases. These reforms had been criticised widely, due to the temporal financial strain it imposes on the least financially resilient people in society. The fact that the exact date of the roll-out varied across the seven metropolitan boroughs within the county allows for a more precise estimation of the effect of UC on the reported number of domestic abuse incidents in the West Midlands. This analyses will constitute the first empirical investigation of the direct effect of welfare policy changes on domestic abuse victimisation.

>>>>

This research programme consists of four research topics aiming to deepen our understanding of domestic abuse victimisation, by exploiting the rich information contained in the CSEW, and combine it with insights from other data sources. First, we will explore the demographic and socio-economic characteristics of victims of domestic abuse and the predictors of suffering serious harm. Second, we will examine the reasons behind not just under-reporting from victims, but also police mis-recording of domestic abuse cases and victim's perception of the police after reporting an incident. Third, we will attempt to quantify the long-term effects of being a victim of domestic abuse and witnessing domestic abuse at home as a child. Finally, we explore the association between exogenous, time-varying factors, such as changes in alcohol consumption and financial resilience, and the prevalence of domestic abuse.

@@

The research projects detailed below draw on various data sources. All of the projects rely on the CSEW, an annual, cross-sectional representative survey collecting information on victims of crime across England and Wales, including 30,000--40,000 households every year \cite{OfficeforNationalStatistics2019}. The survey has a self-completion module containing questions about the respondent's experiences of domestic abuse throughout their life (since the age of 16), and in the past 12 months in particular. Given the serious problem of underreporting, one of the unique characteristics of domestic abuse, the CSEW is the most reliable source of information on the prevalence of domestic abuse in England and Wales. We have access to this dataset via the UK Data Service.

@@

These two datasets have different merits. The CSEW provides us with rich demographic and socio-economic information on the respondent, and to investigate the casual effects of domestic abuse victimisation, we can construct a control group from those individuals who have not experienced domestic abuse. However, it does not contain any information on the perpetrators and has limited longitudinal information. In contrast, the WMP dataset has information on both the victim and the offender, and records all reported incidents for the same victim-offender pair, allowing us to analyse temporal patterns and identify high-risk times of the year. We will not link the individuals in these datasets at any time.

@@

This research programme addresses four overarching topics, exploring different aspects of domestic abuse. Drawing on the extensive, rich respondent information in the CSEW and the unique, temporal nature of the WMP data, this research complements, extends and deepens our understanding of domestic abuse. In outlining the analysis plan for each of these research topics, we will specify a set of statistical analyses we plan to conduct. We will preregister all of our analysis plans and complement all analyses with robustness checks, and further exploratory regressions as the data allows (see, for an example of this approach in some of our \href{https://cog.warwick.ac.uk/~pssmar/Trendl_Stewart_Mullett.pdf}{prior work} on domestic abuse).

@@

Investigating the explanatory power of a broader set of victim characteristics within one (logistic) regression model will provide us with a deeper understanding of the relative importance of these factors in predicting domestic abuse victimisation, compared to a purely descriptive statistics approach. We are also interested in how the predictive power of victim characteristics varies with the type of abuse suffered (physical abuse, threats, sexual abuse).

@@

We will also use the CSEW to explore the police's response to reported cases of domestic abuse. Increasing victim's trust in the Criminal Justice System (CJS) through improving law enforcement response to reported incidents is key for improving willingness to report. Mis-recording of domestic abuse is a serious problem across police forces, for example, a recent inspection reported that WMP have failed to record 25\% of reported crimes that were domestic abuse-related \cite{HerMajestysInspectorateofConstabularyandFires&RescueServices2019}. In this research topic, we will explore whether the extent of crime mis-recording depends on victim characteristics. Using the CSEW, we will be able to tell whether the type of domestic abuse reported by the respondent is likely to amount to a crime according to the list of notifiable offences \cite{countingrules}. A multinomial regression approach will allow us to examine the extent to which police mis-recording is different across broad crime types (sexual, physical violence, harassment and stalking), and depends upon the socio-economic characteristics of the victim and police forces (e.g., funding per police officer, deprivation). Complementing this analysis with a random forest approach will reveal the role of interactions between individual and police force characteristics. Lastly, these analyses can also be applied to examine the effect upon victim satisfaction with the police, and their perception of their personal security after reporting a crime.

@@

 Domestic abuse has long-lasting adverse effects on victims and children witnessing the abuse. In this research topic, our aim is to explore some of these consequences. The 2016 CSEW included a module asking respondents about whether they witnessed domestic abuse at home as a child. Descriptive statistics of this dataset revealed that those who reported to have witnessed abuse as a child were significantly more likely to report long-term health problems, and live in a single parent household as an adult \cite{ONSChildhood}. First, we will consolidate and extend these findings using propensity score matching, a statistical approach that allows for causal inferences to be drawn about the effect of witnessing abuse, whilst also controlling for a range of confounds, such as economic deprivation. 
 %Gaining a deeper understanding of the tangible, far-reaching consequences of domestic abuse will help to design survivor support programmes and quantify the societal harm caused by it.

@@

We will further complement this analysis using WMP data, by identifying whether younger offenders' home addresses can be linked to a past domestic abuse incident. It is not possible to predict whether witnessing domestic abuse increases the chance of becoming a young offender, but we can examine whether it predicts the type of crimes they engage in. For example, whether, after controlling for the socio-economic characteristics of the neighbourhood (using the IMD), such individuals are more likely to commit similarly violent acts than other young offenders.

 These analyses will constitute the first UK-based comprehensive quantitative exploration of the causal effect of domestic abuse and childhood abuse on socio-economic outcomes. The resulting paper will be targeted at The Lancet Public Health, and Social Science and Medicine.

@@

In addition, drawing on previous findings about the link between financial stress and domestic abuse, we will explore what effect the roll-out of the Universal Credit (UC) across the West Midlands had on the reported number of domestic abuse cases. These reforms had been criticised widely, due to the temporal financial strain it imposes on the least financially resilient people in society. The fact that the exact date of the roll-out varied across the seven metropolitan boroughs within the county allows for a more precise estimation of the effect of UC on the reported number of domestic abuse incidents in the West Midlands. These analyses will constitute the first empirical investigation of the direct effect of welfare policy changes on domestic abuse victimisation.
